\chapter{Release Notes}

\section*{Version 1.0.1}

This release is a minor release, with a few optimization features.

\subsection*{Changes to the HDL Core}
\begin{itemize}
\item The multiplier and shifter are now merged together.
\item Added a new ZPU (optional) instruction, FMUL16, opcode (01h), which performs a fixed-point 16:16 multiplication
\item Multiplier is now signed instead of unsigned
\item SPI controller is now able to do 16, 24 and 32-bit transfers
\item UART now has a bit depicting if all transmission has ended. Needed to fix a bug when your sketch changed baudrate and it was still transmitting.
\end{itemize}

\subsection*{IDE/Code}
\begin{itemize}
\item New IDE based on Arduino 1.0.1
\item Updated UARTSTATUS register index
\item IDE now reports a better size for sketches, including program and code size.
\item Removed misinclusion of <new> in some headers
\item IDE can now build and include asssembly (.S) files in the sketch folder
\end{itemize}

\subsection*{Boards}
\begin{itemize}
\item SPI interface is now directly connected on most boards, so to speed up sampling.
\item Add new board (Papilio One 250 with extra 2Kb RAM). All P1 250 are compatible.
\end{itemize}

\subsection*{Libraries}
\begin{itemize}
\item SmallFS was updated to use the new SPI multibyte transfer sizes.
\end{itemize}

\section*{Version 1.0.0}

This release is a major release!. Many many changes were done to all components, and not all of them are
depicted here. Please refer to git changelogs for more detail.

\subsection*{Changes to the HDL Core}

\begin{itemize}
\item New ZPU core - ZPU Extreme!
\item Most cores are now Wishbone compatible.
\item New timer implementation 
\item All IO devices were pushed to the top level module, to ease customization 
\end{itemize}

\subsection*{IDE/Code}

\begin{itemize}
 \item New IDE based on Arduino 1.0
 \item Upload-to-RAM feature (control-click on Upload button!!!)
 \item Code is now smaller due to new toolchain features
 \item Some common functions are now exported by the bootloader, to reduce code size.
 \item Some functions have now assembly implementations for speed.
\end{itemize}

\subsection*{Boards}
\begin{itemize}
 \item New board: Papilio Plus S6LX4
 \item New board: Papilio Plus S6LX9
 \item New board: Nexys2 S3E1200 (thanks to Paul Urbanus)
 \item New board variants!
\end{itemize}

\subsection*{Libraries}
\begin{itemize}
 \item Small fixes here and then.
\end{itemize}

\subsection*{Other/Contributed}
\begin{itemize}
\item SID HDL implementation
\item Pokey HDL implementation
\item YM2149 HDL implementation
\item Audiomixer and simple sigmadelta
\item OpenCores I2C
\end{itemize}

