\section{SigmaDelta DAC}
ZPUino SigmaDelta DAC is a two-channel 16-bit DAC with configurable endianess.
\subsection{HDL sources and modules}

\subsubsection{HDL instantiation template}
\begin{lstlisting}[language=VHDL]
component zpuino_sigmadelta is
  port (
    wb_clk_i: in std_logic;
    wb_rst_i: in std_logic;
    wb_dat_o: out std_logic_vector(wordSize-1 downto 0);
    wb_dat_i: in std_logic_vector(wordSize-1 downto 0);
    wb_adr_i: in std_logic_vector(maxIObit downto minIObit);
    wb_we_i:  in std_logic;
    wb_cyc_i: in std_logic;
    wb_stb_i: in std_logic;
    wb_ack_o: out std_logic;
    wb_inta_o:out std_logic;

    sync_in:  in std_logic;

    -- Connection to GPIO pin
    spp_data: out std_logic_vector(1 downto 0);
    spp_en:   out std_logic_vector(1 downto 0)
  );
end component;
\end{lstlisting}

\subsubsection{Compliance}
The Sigma Delta DAC is wishbone compatible, in non-pipelined mode.

\subsubsection{Source files}
All Sigma Delta system is implemented in \emph{zpuino\_sigmadelta.vhd}.

\subsection{Location}
The Sigma Delta module is usually located in IOSLOT 5.

\subsection{Registers}

\subsubsection{SIGMADELTACTL}
Sigma Delta DAC control register.

\begin{table}[H]
\begin{center}
\begin{tabularx}{14cm}{Xcccc}
31 \hfill 4 & 3 & 2 & 1 & 0 \\

\hline
\multicolumn{1}{|c|}{Reserved} &
\multicolumn{1}{|c|}{SDEXTTRIG}  &
\multicolumn{1}{|c|}{SDLE}  &
\multicolumn{1}{|c|}{SDENA1}  &
\multicolumn{1}{|c|}{SDENA0}  \\

\hline
\end{tabularx}
\caption{SIGMADELTACTL register}
\end{center}
\end{table}

\begin{description}
\item{SDENA0} \hfill \\ SigmaDelta Channel 0 enable
\item{SDENA1} \hfill \\ SigmaDelta Channel 1 enable
\item{SDLE} \hfill \\ Sigma-Delta Little Endian support. If set to 1 input 
is assumed to be little endian, otherwise the native format (big endian) is used.
\item{SDEXTTRIG} \hfill \\ Sigma-Delta external trigger support. If set to 1, then an external signal (Timer0 compare match) is used to transfer previously
written data into the output. This is useful if you require a constant sigma delta update, but you have jitter in the code that do 
not allow for a precise output timing.
\end{description}


\subsubsection{SIGMADELTADATA}
Sigma Delta DAC data register.

\begin{table}[H]
\begin{center}
\begin{tabularx}{14cm}{XX}
31 \hfill 16 & 15 \hfill 0 \\

\hline
\multicolumn{1}{|c|}{SDCH1}  &
\multicolumn{1}{|c|}{SDCH0}  \\

\hline
\end{tabularx}
\caption{SIGMADELTADATA register}
\end{center}
\end{table}

\begin{description}
\item{SDCH0} \hfill \\ SigmaDelta Channel 0 data
\item{SDCH1} \hfill \\ SigmaDelta Channel 1 data
\end{description}



\subsection{Software}
